Figures

### 2016-08-17 data_explore/data_explore.Rmd
incorporating Jamie's comments

``` {r plot_bar_charts_by_taxa_old, echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE}
# ### bar charts comparing number of species in each class for each data set
# 
# #library(colorspace)
# #hclvec <- heat_hcl(3); colvec <- c('AquaMaps' = hclvec[1], 'Both' = hclvec[2], 'IUCN' = hclvec[3])
# #colvec <- c('AquaMaps' = '#1b9e77', 'Both' = '#565080', 'IUCN' = '#d95f02')
# colvec <- c('AquaMaps' = '#76A94C', 'Both' = '#e18258', 'IUCN' = '#9168a9')
# 
# taxa_df_long <- taxa_df %>%
#   dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
#   gather(dataset, n_spp, AquaMaps:Both) %>%
#   group_by(class) %>%
#   mutate(n_total = sum(n_spp, na.rm = TRUE),
#          percent = n_spp/n_total * 100,
#          dataset = factor(dataset, levels = c('AquaMaps', 'Both', 'IUCN'), ordered = TRUE)) %>%
#   arrange(desc(dataset))
# 
# break_nums <- seq(0, 100, 20)
# 
# bar_taxa <- ggplot(taxa_df_long, aes(x = class, y = percent, fill = dataset)) +
#   ggtheme_plot + 
#   theme(panel.grid.major.x = element_blank()) +
#   geom_bar(stat = 'identity', alpha = 1) +
#   geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
#   scale_fill_manual(values = colvec) +
#   geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
#             size = 2, 
#             angle = 90, hjust = 0, color = 'grey30') +
#   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
#   scale_y_continuous(expand = c(0, 0), 
#                      limits = c(0, 130),
#                      breaks = break_nums,
#                      labels = sprintf('%s%%', break_nums)) + 
#   labs(# title = 'Species maps by taxonomic class',
#        x = 'Class', 
#        y = 'Percent of mapped species',
#        fill = 'Dataset')
# 
# print(bar_taxa)
# ggsave(filename = file.path(dir_fig, 'fig1a_bar_by_taxa_all.png'), 
#        width = 17.8, height = 7, units = 'cm', dpi = 600)

@@

taxa_df1 <- taxa_df %>%
  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
  mutate(AquaMaps = ifelse(is.na(AquaMaps), 0, AquaMaps),
         IUCN     = ifelse(is.na(IUCN),     0, IUCN),
         Both     = ifelse(is.na(Both),     0, Both),
         n_total  = AquaMaps + IUCN + Both,
         IUCN_end = 100 * (IUCN + Both)/n_total,
         IUCN_start = 0,
         AquaMaps_end   = 100,
         AquaMaps_start = 100 * (n_total - AquaMaps - Both)/n_total,
         Both_end = AquaMaps_start,
         Both_start = IUCN_end) %>%
  gather(key = dataset,  value = start, ends_with('start')) %>%
  gather(key = dataset2, value = end, ends_with('end')) %>%
  mutate(dataset = str_replace(dataset, '_start', ''),
         dataset2 = str_replace(dataset2, '_end', '')) %>%
  filter(dataset == dataset2) %>%
  select(-dataset2)

taxa_all_df <- taxa_df1 %>%
  group_by(dataset) %>%
  summarize(class = 'Total',
            AquaMaps_total = sum(AquaMaps) + sum(Both),
            IUCN_total = sum(IUCN) + sum(Both),
            n_total  = sum(n_total),
            IUCN_end = 100 * IUCN_total/n_total,
            IUCN_start = 0,
            AquaMaps_end   = 100,
            AquaMaps_start = 100 * (n_total - AquaMaps_total)/n_total,
            Both_end = AquaMaps_start,
            Both_start = IUCN_end) %>%
  select(-dataset) %>% distinct() %>%
  gather(key = dataset,  value = start, ends_with('start')) %>%
  gather(key = dataset2, value = end, ends_with('end')) %>%
  mutate(dataset = str_replace(dataset, '_start', ''),
         dataset2 = str_replace(dataset2, '_end', '')) %>%
  filter(dataset == dataset2) %>%
  select(-dataset2, -AquaMaps_total, -IUCN_total)

taxa_df2 <- taxa_all_df %>%
  bind_rows(taxa_all_df %>%
              mutate(class = ' ',
                     start = 0,
                     end   = 0,
                     n_total = NA)) %>%
  bind_rows(taxa_df1 %>%
              select(-IUCN, -AquaMaps, -Both)) %>%
  mutate(class = factor(class, levels = unique(.$class), ordered = TRUE),
         n_label = ifelse(!is.na(n_total), paste0('n = ', n_total), ''))

@@

bar_taxa <- ggplot(taxa_df2, aes(x = class, color = dataset)) +
  ggtheme_plot + 
  theme(panel.grid.major.x = element_blank()) +
  geom_linerange(aes(ymin = start, ymax = end, size = dataset),
                 position = position_dodge(width = 0.5)) +
  scale_size_manual(values = c('IUCN' = .25, 'AquaMaps' = .25, 'Both' = 2.5)) +
  scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey60', 'Both' = 'black')) +
  geom_text(aes(label = n_label, y = 102),
            size = 2,
            angle = 90, hjust = 0, color = 'grey30') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 130),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums)) +
  labs(x = 'Class',
       y = 'Percent of mapped species',
       fill = 'Dataset')

>>>>

@@

taxa_df1 <- taxa_df %>%
  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Paired = n_class_both) %>%
  mutate(AquaMaps = ifelse(is.na(AquaMaps), 0, AquaMaps),
         IUCN     = ifelse(is.na(IUCN),     0, IUCN),
         Paired     = ifelse(is.na(Paired),     0, Paired),
         n_total  = AquaMaps + IUCN + Paired,
         IUCN_end = 100 * (IUCN + Paired)/n_total,
         IUCN_start = 0,
         AquaMaps_end   = 100,
         AquaMaps_start = 100 * (n_total - AquaMaps - Paired)/n_total,
         Paired_end = AquaMaps_start,
         Paired_start = IUCN_end) %>%
  gather(key = dataset,  value = start, ends_with('start')) %>%
  gather(key = dataset2, value = end, ends_with('end')) %>%
  mutate(dataset = str_replace(dataset, '_start', ''),
         dataset2 = str_replace(dataset2, '_end', '')) %>%
  filter(dataset == dataset2) %>%
  select(-dataset2)

taxa_all_df <- taxa_df1 %>%
  group_by(dataset) %>%
  summarize(class = 'Total',
            AquaMaps_total = sum(AquaMaps) + sum(Paired),
            IUCN_total = sum(IUCN) + sum(Paired),
            n_total  = sum(n_total),
            IUCN_end = 100 * IUCN_total/n_total,
            IUCN_start = 0,
            AquaMaps_end   = 100,
            AquaMaps_start = 100 * (n_total - AquaMaps_total)/n_total,
            Paired_end = AquaMaps_start,
            Paired_start = IUCN_end) %>%
  select(-dataset) %>% distinct() %>%
  gather(key = dataset,  value = start, ends_with('start')) %>%
  gather(key = dataset2, value = end, ends_with('end')) %>%
  mutate(dataset = str_replace(dataset, '_start', ''),
         dataset2 = str_replace(dataset2, '_end', '')) %>%
  filter(dataset == dataset2) %>%
  select(-dataset2, -AquaMaps_total, -IUCN_total)

taxa_df2 <- taxa_all_df %>%
  bind_rows(taxa_all_df %>%
              mutate(class = ' ',
                     start = 0,
                     end   = 0,
                     n_total = NA)) %>%
  bind_rows(taxa_df1 %>%
              select(-IUCN, -AquaMaps, -Paired)) %>%
  mutate(class = factor(class, levels = unique(.$class), ordered = TRUE),
         n_label = ifelse(!is.na(n_total), paste0('n = ', n_total), ''),
         dataset = factor(dataset, levels = c('AquaMaps', 'Paired', 'IUCN'), ordered = TRUE))

@@

bar_taxa <- ggplot(taxa_df2, aes(x = class, color = dataset)) +
  ggtheme_plot + 
  theme(panel.grid.major.x = element_blank()) +
  geom_linerange(aes(ymin = start, ymax = end, size = dataset),
                 position = position_dodge(width = 0.5)) +
  scale_size_manual(values = c('IUCN' = .25, 'AquaMaps' = .25, 'Paired' = 2.5)) +
  scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey60', 'Paired' = 'black')) +
  geom_text(aes(label = n_label, y = 102),
            size = 2,
            angle = 90, hjust = 0, color = 'grey30') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 130),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums)) +
  labs(x = 'Class',
       y = 'Percent of mapped species',
       fill = 'Dataset')

### 2016-08-17 ms_3_res_disc_concl.Rmd
incorporating Jamie's comments

Neither IUCN nor AquaMaps claims a lock on the "truth", and the small overlap in mapped species means each provides value in taxonomic and geographic coverage.  Method-driven differences in commission and omission errors produce clear and significant disagreement in species range descriptions between AquaMaps and IUCN datasets.  To examine the implications of these differences, we repeated two recent studies, varying only the selection of one data set over the other.

>>>>

__Neither IUCN nor AquaMaps claims a lock on the "truth", and the small overlap in mapped species means each provides value in taxonomic and geographic coverage.  Method-driven differences in commission and omission errors produce clear and significant disagreement in species range descriptions between AquaMaps and IUCN datasets.__  To examine the implications of these differences, we repeated two recent studies, varying only the selection of one data set over the other.
